DOM Review: Initial analysis

### 2023-02-03 Analysis/Code/2_InitialAnalysis.Rmd
added some obj plots

## Map of authorship countries by # of papers
studyData %>%
  # Get count of 
  group_by(`First author country`) %>%
  summarise(Count = n()) %>%
  arrange(-Count) %>% 
  mutate(`First author country` = case_when(`First author country` == "United States of America" ~ "United States",
                                            `First author country` != "United States of America" ~ `First author country`)) %>%
  # Merge by country name
  right_join(countriesShapefile, by = c('First author country' = 'COUNTRY')) %>%
  # Plot map
  ggplot() +
  geom_sf(aes(geometry = geometry, fill = log(Count))) +
  labs(title = "Log paper count by country") +
  theme_void() + 
  scale_fill_continuous(name = "Log counts", low="thistle2", high="darkred",na.value="white")

@@

```{r Affiliation types, echo = FALSE}
studyData %>%
  summarise(NGO = sum(Author_NGO)/n()*100,
            Academic = sum(Author_Academic)/n()*100,
            Government = sum(Author_Government/n()*100),
            Private = sum(Author_Private)/n()*100) %>% 
  pivot_longer(cols = c(NGO, Academic, Government, Private), names_to = 'Affiliation', values_to = "% of papers") %>%
  arrange(-`% of papers`) %>%
  kable(digits = 1)
```

## Objectives
Five categories of objectives and goals were assessed for each paper:
* Assessing trends - interested in the trajector of estimated or derived parameters
* Testing hypotheses - evaluating attributes of a **specific** covariate on parameters
* Spatial mapping - Extrapolating estimates of papameter values beyond surveyed sites
* Future projections - Making predictions of parameters into the future
* Methods - introducing, testing, or demonstrating a new extension to these models

@@

```{r Taxa, echo = FALSE}

@@

@@

@@

>>>>

## Map of authorship countries by # of papers
studyData %>%
  # Get count of papers by author's first country
  group_by(`First author country`) %>%
  summarise(Count = n()) %>%
  arrange(-Count) %>% 
  mutate(`First author country` = case_when(`First author country` == "United States of America" ~ "United States",
                                            `First author country` != "United States of America" ~ `First author country`)) %>%
  # Merge by country name
  right_join(countriesShapefile, by = c('First author country' = 'COUNTRY')) %>%
  # Plot map
  ggplot() +
  geom_sf(aes(geometry = geometry, fill = log(Count))) +
  labs(title = "Log paper count by country") +
  theme_void() + 
  scale_fill_continuous(name = "Log counts", low="thistle2", high="darkred",na.value="white")

@@

```{r Affiliation types, echo = FALSE}
# Calc % of each type
studyData %>%
  summarise(NGO = sum(Author_NGO)/n()*100,
            Academic = sum(Author_Academic)/n()*100,
            Government = sum(Author_Government/n()*100),
            Private = sum(Author_Private)/n()*100) %>% 
  # Make legible for table
  pivot_longer(cols = c(NGO, Academic, Government, Private), names_to = 'Affiliation', values_to = "% of papers") %>%
  arrange(-`% of papers`) %>%
  kable(digits = 1)
```

## Objectives
Five categories of objectives and goals were assessed for each paper:

* Assessing trends - interested in the trajector of estimated or derived parameters
* Testing hypotheses - evaluating attributes of a **specific** covariate on parameters
* Spatial mapping - Extrapolating estimates of papameter values beyond surveyed sites
* Future projections - Making predictions of parameters into the future
* Methods - introducing, testing, or demonstrating a new extension to these models

@@

Monitoring trends and testing hypotheses were the most popular objectives; making predictions (spatial or temporal) were quite rare.
```{r Objectives table, echo = FALSE}
# Make table 
objData %>% 
  pivot_longer(cols = c(Obj_Methods, Obj_Projections, Obj_Hypothesis, Obj_Spatial, Obj_Trends),
               names_to = "Objective", values_to = "TF") %>%
  # Drop obj_ prefix
  mutate(Objective = str_replace(Objective, "Obj_","")) %>%
  # Tally up and make table
  group_by(Objective) %>%
  summarise(`Percent of papers` = 100*sum(TF)/54) %>%
  arrange(-`Percent of papers`) %>%
  kable(digits = 1)
```

```{r Obj by Authorship group, echo = FALSe}
objByAuth <- full_join(objData, studyData, by = c("Review ID", "Title")) %>%
  # Pivot authorship column and adjust  
  pivot_longer(cols = c(Author_NGO, Author_Academic, Author_Government, Author_Private), 
               names_to = 'Authorship', values_to = "Auth_TF") %>%
  mutate(Authorship = str_replace(Authorship, "Author_","")) %>%
  # Get count of # of papers by authorship for the plot later
  filter(Auth_TF == TRUE) %>%
  group_by(Authorship) %>%
  mutate(nAuth = n()) %>% 
  # Pivot objectives column and adjust
  pivot_longer(cols = c(Obj_Methods, Obj_Projections, Obj_Hypothesis, Obj_Spatial, Obj_Trends),
               names_to = "Objective", values_to = "Obj_TF") %>%
  mutate(Objective = str_replace(Objective, "Obj_","")) %>%
  filter(Obj_TF == TRUE)

objByAuth %>%
  filter(Obj_TF == TRUE & Auth_TF == TRUE) %>%
  group_by(Authorship) %>%
  mutate(TotalAuth = length(unique(Title))) %>%
  group_by(Authorship, Objective, nAuth) %>%
  summarise(Count = n()/nAuth) %>%
  ungroup() %>%
  complete(Authorship, Objective) %>%
  mutate(Count = case_when(is.na(Count) ~ 0.01,
                           !is.na(Count) ~ Count)) %>%

  ggplot(aes(x = Authorship, y = Count, fill = Objective)) +
  geom_col(position = "dodge") +
  ylim(0, 1) + 
  labs(title = "Percent of papers by affiliation pursuing each objective")
```

```{r Covs by obj, echo = FALSE}
covData <- reviewResults[["Covariates"]] %>%
  mutate(`Review ID` = as.numeric(str_replace(`Review ID`, "[:alpha:]", "")))

objData %>%
   pivot_longer(cols = c(Obj_Methods, Obj_Projections, Obj_Hypothesis, Obj_Spatial, Obj_Trends),
               names_to = "Objective", values_to = "Obj_TF") %>%
  mutate(Objective = str_replace(Objective, "Obj_","")) %>%
  filter(Obj_TF == TRUE) %>%
  select(`Review ID`, Objective) %>%
  full_join(covData, by = 'Review ID') %>%

  group_by(`Review ID`, Objective, Parameter) %>%
  summarise(Count = n()) %>%
  filter(Parameter %in% c("Colonisation", "Extinction", "InitialOccupancy", "Detection"),
         Count < 20) %>%

  ggplot(aes(x = Objective, y = Count, colour = Parameter)) +
  geom_boxplot()

@@

## Taxa and location
Details on the study organism was collected for each study to assess any trends in the types of species these models are applied to, with a particular focus on those that are vulnerable to conservation threats. The locations where surveys were conducted was also noted.

```{r Taxa, echo = FALSE}
taxaData <- reviewResults[["Taxa"]]
```

As with authorship, the United States dominated the countries where surveys took place.

### Study site maps {.tabset}
#### Country
```{r Site country map, echo = FALSE}
taxaData %>%
  # Divide by unique title/country
  group_by(Country, Title) %>%
  summarise() %>%
  # Get count of papers by site country
  group_by(Country) %>%
  summarise(Count = n()) %>%
  arrange(-Count) %>% 
  mutate(Country = case_when(Country == "United States of America" ~ "United States",
                             Country != "United States of America" ~ Country)) %>%
  # Merge by country name
  right_join(countriesShapefile, by = c('Country' = 'COUNTRY')) %>%
  # Plot map
  ggplot() +
  geom_sf(aes(geometry = geometry, fill = log(Count))) +
  labs(title = "Log paper count by country") +
  theme_void() + 
  scale_fill_continuous(name = "Log counts", 
                        low = "thistle2", 
                        high = "darkred",
                        na.value = "white")

# Table with the same data
taxaData %>%
  # Divide by unique title/country
  group_by(Country, Title) %>%
  summarise() %>%
  # Get count of papers by site country
  group_by(Country) %>%
  summarise(`Total papers` = n()) %>%
  arrange(-`Total papers`) %>% 
  kable()
```

#### Continent
```{r Site continent map, echo = FALSE}
continentShapefile <- st_read("Analysis/SpatialData/Continents/World_Continents.shp")

taxaData %>%
  # Divide by unique title/country
  group_by(Country, Continent, Title) %>%
  summarise() %>%
  # Get count of papers by site continent
  group_by(Continent) %>%
  summarise(Count = n()) %>%
  arrange(-Count) %>% 
  mutate(Continent = case_when(Continent == "Oceania" ~ "Australia",
                               Continent != "Oceania" ~ Continent)) %>%

  # Merge by continent name
  right_join(continentShapefile, by = c('Continent' = 'CONTINENT')) %>%
  # Plot map
  ggplot() +
  geom_sf(aes(geometry = geometry, fill = Count)) +
  labs(title = "Paper count by continent") +
  theme_void() + 
  scale_fill_continuous(name = "Log counts", 
                        low = "thistle2", 
                        high = "darkred",
                        na.value = "white")

# Table with the same data
taxaData %>%
  # Divide by unique title/country
  group_by(Country, Title, Continent) %>%
  summarise() %>%
  # Get count of papers by continent
  group_by(Continent) %>%
  summarise(`Total papers` = n()) %>%
  arrange(-`Total papers`) %>% 
  kable()

#### Realm
```{r Site realm map, echo = FALSE}
#ecoregionsShapefile <- st_read("Analysis/SpatialData/Ecoregions/Terrestrial_Ecoregions.shp") %>%
#  st_make_valid() 
#
#t <- ecoregionsShapefile %>%
#  group_by(WWF_REALM2) %>%
#  summarise(geometry = st_combine(geometry))
#
#taxaData %>%
#  # Divide by unique title/country
#  group_by(Country, Realm, Title) %>%
#  summarise() %>%
#  # Get count of papers by site continent
#  group_by(Realm) %>%
#  summarise(Count = n()) %>%
#  arrange(-Count) %>% 
#  
#  # Merge by realm
#  right_join(ecoregionsShapefile, by = c('Realm' = 'WWF_REALM2')) %>%
#  # Plot map
#  ggplot() +
#  geom_sf(aes(geometry = geometry, fill = Count)) +
#  labs(title = "Paper count by Udvardy realm") +
#  theme_void() + 
#  scale_fill_continuous(name = "Log counts", 
#                        low = "thistle2",
#                        high = "darkred",
#                        na.value="white")
#
# Table with the same data
taxaData %>%
  # Divide by unique title/country
  group_by(Country, Title, Realm) %>%
  summarise() %>%
  # Get count of papers by site realm
  group_by(Realm) %>%
  summarise(`Total papers` = n()) %>%
  arrange(-`Total papers`) %>% 
  kable()
```

## Survey methods

@@

## Parameters and covariates

@@

## Modelling decisions
